
## Integrative cluster analysis

```{r setup, include =FALSE,echo = FALSE,eval=TRUE}

rm(list = ls())
graphics.off()

library(ggplot2)
library(data.table)
library(RColorBrewer)
library(reshape2)
library(ggdendro)
library(broom)
library(grid)
library(gridExtra)
library(cluster)
library(GenomicRanges)
library(gtable)

## Analysis parameters
K <- 10  # number of clusters
top <- 10 # number of top peaks per cluster
cutline <- .75 # minimal proportion of TFs regions overlaping with peaks in cluster


library(knitr)
opts_knit$set(root.dir=normalizePath('../'))
opts_chunk$set(fig.path = "../figures/cluster_integrative/")

## loads binary matrices + functions to generate them
load("../data/ranges/all_EBV_GenomicRanges.RData")
load("../data/RData/pam_clusters.RData")

source("../R/heatmap_functions.R")
source("../R/clustering_analysis.R")

peaks = c("EBNA2","EBNA3A","EBNA3B","EBNA3C","RBPJ")
pattern = c("EBNA","RBPJ","H?K","Dnase","H2Z")

cols_to_remove <- function(patterns,set)
{
  ll = lapply(pattern,FUN = grepl_cmd,set)
  n = length(ll)
  ll = as.list(paste0(ll,c(rep("&",n-1),"")))
  return(do.call(paste0,ll))
}

get_names <- function(range_data) return(names(range_data@elementMetadata@listData))

grepl_cmd <- function(variable,set)
{
return(paste0("!grepl(",change_to_str_val(variable),",",set,")",sep =""))
}

change_to_str_val <- function(str) return(paste0("'",str,"'"))

mc = 8

mats <- mcmapply(build_binary_matrix, names(ranges),ranges,
  MoreArgs = list(expr = "Dnase == 1",
             col_expr = cols_to_remove(pattern,'columns')),
			 SIMPLIFY=FALSE,mc.silent=TRUE,mc.cores = mc)

mats <- mats[peaks]

```

In the previous analysis we considered clustering the peak set
specific matrices filtering the regions that didn't overlap DHS and
then we calculated the frequency of overlapping peaks in a cluster
with a specific TF. That way, we considered two strategies to
integrate the clusters of the different peak sets.


```{r cluster_prop,include=FALSE,echo=FALSE,eval=TRUE}
pam_clust <- mclapply(mats,pam,K,mc.cores = mc,mc.preschedule = TRUE)

mats_dt <- lapply(mats,data.table)
mats_dt <- mapply(function(x,y){
  x[,clustering:=y$clustering]
  return(x)},mats_dt,pam_clust,SIMPLIFY=FALSE)
melt_mats <- lapply(mats_dt,melt,id.vars="clustering")

melt_mat <- mapply(function(x,y){
  x[,set:=y]
  return(x)},melt_mats,names(melt_mats),SIMPLIFY=FALSE)
melt_mat <- do.call(rbind,melt_mat)

props <- melt_mat[,mean(value), by=.(clustering,variable,set)]
lengths <- melt_mat[,length(value),by=.(clustering,variable,set)][,(V1)]

props[,group:=paste(lengths,"peaks")]
  
top_prop <- props[,top_k(V1,variable,top),by=.(set,clustering)]
setnames(top_prop,names(top_prop),c("set","clustering","prop","variable"))
cut_prop <- props[V1 >= cutline]
setnames(cut_prop,names(cut_prop),c("clustering","variable","set","prop","group"))

```

```{r incidence_mats,include=FALSE,echo=FALSE,eval=TRUE}

top_mat <- build_integrative_mat(top_prop)
cut_mat <- build_integrative_mat(cut_prop)
top_hm <- gg_heatmap(top_mat,row.names=TRUE)
cut_hm <- gg_heatmap(cut_mat,row.names=TRUE)

```

That way we can build two different incidence matrices.

1. One where the rows represent one of the `r K` clusters of each
   dataset (i.e. `r length(peaks)*K` rows) and the columns are the
   union of the top `r top` TFs for each peaksets.
  
2. Again the rows represent one of the `r K` clusters of each datasets
   (excepting the clusters where there are not TFs such that its
   overlap frequency is greater `r cutline*100`%). And the columns
   represent the union of all TFs such that its overlapping frequency
   respect the peaks in the cluster is greater than `r cutline*100`%.

```{r clust_funs,include=FALSE,echo=FALSE,eval=TRUE}
clust_plot <- function(data,clust)
{
dt <- data[clustering == clust]
p <- ggplot(dt,aes(reorder(variable,-V1),V1))+
  geom_bar(stat = "identity",width=.8)+
  theme(axis.text.x = element_text(angle = 90,size = 6))+xlab("TF")+ylab("Cluster prop.")+
  facet_wrap(~group)+scale_y_continuous(limits = c(0,1))

return(p)
  
}

arrange_plot <- function(pp,nrow,set)
{
  args.list <- c(pp,nrow,set)
  names(args.list) <- c(paste0("p",1:length(pp)),"nrow","main")
  do.call(grid.arrange,args.list)
}

```


   
### Using the union of the top `r top` TF per cluster as columns

```{r hm_top ,fig.width=6,fig.heigth=8,dpi=120,echo=FALSE}

print_hm(top_hm)

```

- This heatmap divides the drivers as:
  * EBF1, FOXM1, P300, BCL11A, PAX5, RUNX3, NFIC and ATF2
  * POL2, CTCF, PU1
  * TAF1, PML, MXI1, CHD2, ELF1, MAZ, YY1
  * IRF4, BATF, MTA3, MEF2A

Roughly we can see the rows are partitioned into four groups,
therefore we can see the cluster presence frequency for the columns of
the matrix:

```{r top_clusters,include=FALSE,echo=FALSE,eval=TRUE}

p <- 4
clusters <- cutree(hclust(dist(top_mat)),k=p)
top_mat <- data.table(top_mat)
top_mat[,clustering:= clusters]
top_mat <- melt(top_mat,id.vars = "clustering")

top_props <- top_mat[,mean(value),by=.(clustering,variable)]
top_lengths <- top_mat[,length(value),by=.(clustering,variable)]
top_props[,lengths:=top_lengths[,(V1)]]
top_props[,group:=paste(lengths,"rows")]

rm(top_lengths)
top_props[,lengths:=NULL]

top_plots <- lapply(1:p,function(j,data)clust_plot(data,j),top_props)

```

```{r top_propotions,include=TRUE,echo=FALSE,fig.width=16,fig.heigth=12,dpi=1000}
nrow = 2
arrange_plot(top_plots,nrow,paste("Top",top," peaks strategy"))
```



  

### Using the union of the TFs with clustered frequency greater than `r cutline*100`

  
   
```{r hm_cut ,fig.width=6,fig.heigth=8,dpi=120,echo=FALSE}

print_hm(cut_hm)

```

- This one divides the drivers as:
  * EBF1, RUNX3, PAX5, BCL11A, ATF2, NFIC, BATF, P300
  * MAX, TAF1, ELF1, POL2, YY1, MAZ, MXI1
  * SP1, PML, NFATC1, TBLR1, MEF2A, BCL3, MTA3

In this heatmaps we can see that the rows are divided into 3 segments:
  
```{r cut_clusters,include=FALSE,echo=FALSE,eval=TRUE}

p <- 3
clusters <- cutree(hclust(dist(cut_mat)),k=p)
cut_mat <- data.table(cut_mat)
cut_mat[,clustering:= clusters]
cut_mat <- melt(cut_mat,id.vars = "clustering")

cut_props <- cut_mat[,mean(value),by=.(clustering,variable)]
cut_lengths <- cut_mat[,length(value),by=.(clustering,variable)]
cut_props[,lengths:=cut_lengths[,(V1)]]
cut_props[,group:=paste(lengths,"rows")]

rm(cut_lengths)
cut_props[,lengths:=NULL]

cut_plots <- lapply(1:p,function(j,data)clust_plot(data,j),top_props)

```

```{r cut_propotions,include=TRUE,echo=FALSE,fig.width=16,fig.heigth=12,dpi=1000}
nrow = 2
arrange_plot(cut_plots,nrow,paste("Cut",cutline*100,"% peaks strategy"))
```



