

```{r init,include =FALSE,echo = FALSE,eval = TRUE}

library(data.table)
library(knitr)

```

## Association tests among peak sets


Using the annotation into those three categories, we performed
chi-square tests to compare the annotation distribution among the
different sets.

```{r assoc,include=FALSE,echo=FALSE,eval=TRUE}

load("../data/RData/association_tests.RData" )

assoc <- out[["assoc"]]
gof <- out[["gof"]]

assoc[,parameter := NULL]
assoc <- assoc[ !grepl("JK",Set1) & !grepl("JK",Set2)]

gof[,parameter := NULL]
gof <- gof[  !grepl("JK",Set1) & !grepl("JK",Set2)]

gof <- gof[ Set1 != Set2]

```

```{r assoc2,include=TRUE,eval=TRUE,echo=FALSE,results='markup'}

kable(assoc,format = 'markdown',digits = 4)

```

Here we can see that almost all pairs are strongly associated on this
partition (promoters,gene body and intergenic), except EBNA2 and
EBNA3B.


```{r gof,include =TRUE,eval= TRUE,echo =FALSE,results = 'markup'}

kable(gof,format = 'markdown',digits = 4)

```

This are goodness of fit tests, therefore we are testing if the
probability distributions of the first set could be generated as a
sample from the probability distribution of the second set.

## Conditional on Dnase

```{r assoc_dnase,include=FALSE,echo=FALSE,eval=TRUE}

load("../data/RData/association_tests_dnase.RData" )

assoc <- out[["assoc"]]
gof <- out[["gof"]]

assoc[,parameter := NULL]
assoc <- assoc[ !grepl("JK",Set1) & !grepl("JK",Set2)]

gof[,parameter := NULL]
gof <- gof[  !grepl("JK",Set1) & !grepl("JK",Set2)]

gof <- gof[ Set1 != Set2]

```

```{r assoc_dnase2,include=TRUE,eval=TRUE,echo=FALSE,results='markup'}

kable(assoc,format = 'markdown',digits = 4)

```

Again in thise case, we are seeing that almost all pairs of peak
sets are strongly associated conditional on Dnase binding except for
EBNA2 and EBNA3B again.


```{r gof_dnase,include =TRUE,eval= TRUE,echo =FALSE,results = 'markup'}

kable(gof,format = 'markdown',digits = 4)

```

Similarly this are goodnes of fit tests as above where we are testing
if the conditional distribution on Dnase binding could be generated
from the second set's distribution.


Note: This analysis doesn't consider any spatial information about the
peak sets.

