
## Enhancer segmentation of EBNA peaks

For this part, we considered the annotation of the EBNA to ENCODE's
enhancer segmentation. This segmentation is the latest, and the states
considered are explained in
[here](https://genome.ucsc.edu/cgi-bin/hgTrackUi?g=wgEncodeAwgSegmentation&db=hg19):

- TSS	Predicted promoter region including TSS
- PF	Predicted promoter flanking region
- E		Predicted enhancer
- WE	Predicted weak enhancer or open chromatin cis regulatory element
- CTCF	CTCF enriched element
- T		Predicted transcribed region
- R		Repressed or Low Activity region




The data used for our case was extracted from
[here](https://www.encodeproject.org/datasets/ENCSR212BHV/)

Roughly speaking, we took all the EBV peaks (considering two separate
cases, before and after filtering the ones that overlap Dnase
hypersensitive sites), and look for annotation that corresponds
according to the chromHMM - Segway segmentation. In case of a tie, we
considered the annotation as the one that occupied the largest number
of bp in the peak. We have built the chance to use the first or last
labels too respect to the 5' coordinates.

```{r init1,include= FALSE , echo = FALSE, eval = TRUE}


library(knitr)
opts_knit$set(root.dir=normalizePath('../'))
opts_chunk$set(fig.path = "../figures/chromHMM_Segway/")

library(data.table)
library(RColorBrewer)
library(ggplot2)

```

### All the peaks

```{r allpeaks,include = FALSE,eval = TRUE,echo = FALSE}

load("data/RData/chromHMM_segway_proportions.RData")
tab <- chromHMM[[1]]
plot <- chromHMM[[3]]


```

The number of peaks that overlap each annotation are given by:

```{r annot,include= TRUE,echo = FALSE,eval = TRUE}
tab

```

```{r barplot,include = TRUE,echo = FALSE,eval = TRUE, fig.width = 7,fig.height = 5,dpi=200}
plot
```

### Overlap with DHS

```{r dnase,include = FALSE,eval = TRUE,echo = FALSE}

load("data/RData/chromHMM_segway_proportions_dnase.RData")
tab <- chromHMM[[1]]
plot <- chromHMM[[3]]


```

The number of peaks that overlap each annotation and DHS are given by:

```{r annot2,include= TRUE,echo = FALSE,eval = TRUE}
tab

```

```{r barplotDnase,include = TRUE,echo = FALSE,eval = TRUE, fig.width = 7,fig.height = 5,dpi=200}
plot
```

