
## Enhancer segmentation of EBNA peaks

For this part, we considered the annotation of the EBNA to ENCODE's
enhancer segmentation. This segmentation is older and was made by
using the hg18 genome, therefore there are some peaks for which there
is no state available. The states are explained in
[here](https://genome.ucsc.edu/cgi-bin/hgTrackUi?db=hg19&g=wgEncodeBroadHmm)

Roughly speaking, we took all the EBV peaks (considering two separate
cases, before and after filtering the ones that overlap Dnase
hypersensitive sites), and look for annotation that corresponds
according to the liftOver'ed chromHMM segmentation (from hg18 to
hg19). In case of a tie, we considered the annotation as the one that
occupied the largest number of bp in the peak. We have built the
chance to use the first or last labels too respect to the 5'
coordinates.

Since in the original annotation there are 3 labels that are repeated,
those were joined together with their respective pair and the
enumeration was altered a little bit. In total we had 13 categories:

- Original 15
- Minus 3 repeated
- Plus 1 added with the regions missing by the liftOver


```{r init1,include= FALSE , echo = FALSE, eval = TRUE}


library(knitr)
opts_knit$set(root.dir=normalizePath('../'))
opts_chunk$set(fig.path = "../figures/chromHMM/")

library(data.table)
library(RColorBrewer)
library(ggplot2)

getPalette1 <- colorRampPalette(brewer.pal(8,"Set1"))
getPalette2 <- colorRampPalette(brewer.pal(8,"Set2"))
vals <- c( getPalette1(6), getPalette2(7))

```

### All the peaks

```{r allpeaks,include = FALSE,eval = TRUE,echo = FALSE}

load("data/RData/chromHMM_proportions.RData")
tab <- chromHMM[[1]]
plot <- chromHMM[[3]]
pies <- chromHMM[[4]]

load("data/RData/chromHMM_proportions_dnase.RData")
tab2 <- chromHMM[[1]]
plot2 <- chromHMM[[3]]
pies2 <- chromHMM[[4]]

labs <- colnames(tab)

```

The number of peaks that overlap each annotation are given by:

```{r annot,include= TRUE,echo = FALSE,eval = TRUE}
tab

```

```{r barplot,include = TRUE,echo = FALSE,eval = TRUE, fig.width = 8,fig.height = 6,dpi=600,warning = FALSE}

plot+theme(axis.text.y = element_text(size = 5))

```

```{r pies,include =TRUE,echo = FALSE,eval = TRUE, fig.width = 5, fig.height = 5 ,dpi = 500,warning = FALSE}

u <-lapply(pies,print)

```

### Overlap with DHS

The number of peaks that overlap each annotation and DHS are given by:

```{r annot2,include= TRUE,echo = FALSE,eval = TRUE}
tab2

```

```{r barplotDnase,include = TRUE,echo = FALSE,eval = TRUE, fig.width = 8,fig.height = 6,dpi=600,warning = FALSE}

plot2+theme(axis.text.y = element_text(size = 5))

```

```{r piesDnase,include =TRUE,echo = FALSE,eval = TRUE, fig.width = 5, fig.height =5 ,dpi = 500,warning = FALSE}

u <-lapply(pies2,print)

```
