
## Enhancer segmentation of EBNA peaks

For this part, we considered the annotation of the EBNA to ENCODE's
enhancer segmentation. This segmentation is older and was made by
using the hg18 genome, therefore there are some peaks for which there
is no state available. The states are explained in
[here](https://genome.ucsc.edu/cgi-bin/hgTrackUi?db=hg19&g=wgEncodeBroadHmm)

Roughly speaking, we took all the EBV peaks (considering two separate
cases, before and after filtering the ones that overlap Dnase
hypersensitive sites), and look for annotation that corresponds
according to the chromHMM - Segway segmentation. In case of a tie, we
considered the annotation as the one that occupied the largest number
of bp in the peak. We have built the chance to use the first or last
labels too respect to the 5' coordinates.

```{r init1,include= FALSE , echo = FALSE, eval = TRUE}


library(knitr)
opts_knit$set(root.dir=normalizePath('../'))
opts_chunk$set(fig.path = "../figures/chromHMM/")

library(data.table)
library(RColorBrewer)
library(ggplot2)

getPalette <- colorRampPalette(brewer.pal(9,"Set1"))


```

### All the peaks

```{r allpeaks,include = FALSE,eval = TRUE,echo = FALSE}

load("data/RData/chromHMM_proportions.RData")
tab <- chromHMM[[1]]
plot <- chromHMM[[3]]

load("data/RData/chromHMM_proportions_dnase.RData")
tab2 <- chromHMM[[1]]
plot2 <- chromHMM[[3]]

```

```{r test}
identical(plot,plot2)
```


The number of peaks that overlap each annotation are given by:

```{r annot,include= TRUE,echo = FALSE,eval = TRUE}
tab

```

```{r barplot,include = TRUE,echo = FALSE,eval = TRUE, fig.width = 8,fig.height = 6,dpi=500,warning = FALSE}
plot+theme(axis.text.y = element_text(size = 5))
```

### Overlap with DHS

The number of peaks that overlap each annotation and DHS are given by:

```{r annot2,include= TRUE,echo = FALSE,eval = TRUE}
tab2

```

```{r barplotDnase,include = TRUE,echo = FALSE,eval = TRUE, fig.width = 8,fig.height = 6,dpi=500,warning = FALSE}
plot2+theme(axis.text.y = element_text(size = 5))
```

